---
output:
  pdf_document: default
  html_document: default
---
 ---
title: "R Notebook"
output: html_notebook
---

This data set consists of 427,323 records and 13 variables, including the day of birth according to the month and the day of week (DOB_MM, DOB_WK) and other variables like the estimated gestation age in weeks (ESTGEST). We will analyze this data in order to to unearth some information / get to know the data by doing some visualizations.

Let's glimpse at the first 5 rows of our data to get a sens of our data

```{r}
library(lattice)
library(tidyverse)
library(nutshell)
data(births2006.smpl)
birth2006<- births2006.smpl
dim(birth2006)
birth2006[1:5,]
```

## We will look at the frequency of births per day of the week .  

This below graphs will help us understand the distribution of our data; We will take a look at the distribution of date of birth in order to know in which day of the week there were more birth: We see that most the women gave birth between 3rd and 5th days of the week. For instance in the 3rd (Wednesday), 4th (Thursday) and 5th (Friday) , there were almost 7000 birth

```{r}
library(ggplot2)

ggplot(birth2006, aes(DOB_WK ,fill =SEX)) +
  geom_bar()

```

As we can see Thursday has the highest number of birth with a record of *70290* while Monday has the lowest birth with a record of *40274*

```{r}
birth_dow<-table(birth2006$DOB_WK)
birth_dow
```
```{r}
barplot(birth_dow , ylab="Day of Week")

```


## Birth Delivery method:

We will classify the ways of birth delivery according to the  the week and the method of the delivery

```{r}
dob_tbl<-table(WK=birth2006$DOB_WK, MM=birth2006$DMETH_REC)
dob_tbl 
```

As we can see there, vaginal delivery method was the method that was mostly used in 2006.

```{r}
dob_tbl_f <- as.data.frame.matrix(dob_tbl, stringsAsFactors = default.stringsAsFactors())

barchart(dob_tbl[,-2],horizontal=FALSE,groups=FALSE,xlab="Day of Week")
```

```{r}
histogram(~DBWT|DMETH_REC, data = birth2006, layout=c(1,3))

```

The above bar chart of the frequencies of births according to the day of week of the birth 
shows that fewer births take place during the weekend (days 1=Sunday, 2=Monday, ...,7=Saturday of DOB_WK).
This may have to do with the fact that many babies are delivered by cesarean section, 
and  those deliveries are usually scheduled during the week and not on weekends

```{r}
histogram(~DBWT|DPLURAL, data=birth2006, layout=c(1,5), plot.points=FALSE,col="black")

```

```{r}
densityplot(~DBWT|DPLURAL, data=birth2006,layout=c(1,5), plot.points=FALSE, col="black")
densityplot(~DBWT, groups = DPLURAL, data=birth2006, plot.points=FALSE, size=3)
dotplot(~DBWT|DPLURAL, data=birth2006, layout=c(1,5), plot.points=FALSE,col="black")
```

This Scatter plots is shown for birth weight against weight gain: We see that single & Twin birth babies' weight the most.
In addition, the distribution the single birth is normal, Twin is almost normal too; but  triplet, quadruplet, and quintuplet distributions are left skewed, probably those babies  don't weight as the single & Twin babies.  We can conclude that twin, triplet, quadruplet, and quintuplet birht have lower birth rates

```{r}
#xyplot(DBWT~WTGAIN, data=birth2006)
xyplot(DBWT~WTGAIN|DPLURAL, data = birth2006, layout=c(1,5))
#xyplot(DBWT~WTGAIN|DPLURAL, data=birth2006, layout=c(1,5))
```


In this section we will Calculate the avarage birth weight for males & females

```{r}
fac<-factor(birth2006$DPLURAL)
res<-birth2006$DBWT

avg_Per_Gender<- tapply(birth2006$DBWT, INDEX = list(birth2006$DPLURAL, birth2006$SEX), FUN = mean, na.rm=T)
avg_Per_Gender
```

 Here we will finally Calculate the avarage birth weight for males & females

This below graph illustrate the conclusion we made earlier. Single babies have almost twice the weight of Triplet.

```{r}
fac<-factor(birth2006$DPLURAL)
res<-birth2006$DBWT
avg_Per_Birth<-tapply(res, fac, mean, na.rm=T)
avg_Per_Birth
barplot(avg_Per_Birth, ylab="DBWT")

```

Frequency distribution table of estimated gestation period indicates that "99" is the code for "UNKNOW".
The following calculations will omit all records with unknown gestation period(vaue 99)



```{r}
avg_Per_Gender<-table(birth2006$ESTGEST)
new<-birth2006[birth2006$ESTGEST != 99,]
GestTable<- table(new$ESTGEST)
GestTable
```
